---

Applied Time Series Analysis

ADS 506 , Final Project
 
Nicholas	Lee
Hoori Javadnia

Unemployment Rate in California (CAUR)

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE)
```
### libraries


```{r}
library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('grid') # visualisation
library('gridExtra') # visualisation
library('corrplot') # visualisation
library('ggrepel') # visualisation
library('RColorBrewer') # visualisation
library('data.table') # data manipulation
library('dplyr') # data manipulation
library('readr') # data input
library('tibble') # data wrangling
library('tidyr') # data wrangling
library('lazyeval') # data wrangling
library('broom') # data wrangling
library('stringr') # string manipulation
library('purrr') # string manipulation
library('forcats') # factor manipulation
library('lubridate') # date and time
library('forecast') # time series analysis
library('prophet') # time series analysis
```

```{r}
library(tidyverse)
library(MASS)
library(fpp2)
library(corrplot) #for visualisation of correlation
library(lattice) #for visualisation
library(ggplot2) #for visualisation
library(moments)
library(caTools) #for splittind data into testing and training data
library(dplyr) #manipulating dataframe

```



### Data loading and cleaning`

```{r}

data <- read.csv("/Users/javadniahoori/Downloads/CAUR.csv")
head(data,10)
dim(data)
str(data)# Take a look at the class of the dataset
summary(data)
```

## EDA(summary statistics and plots in R)

### missing values
```{r}
sum(is.na(data)) # no missing value
```

```{r}

caur.ts <- ts(data$CAUR, start = c(1976,1,1), end = c(2022,9,1), freq = 12)
plot(caur.ts, xlab = "date", ylab = "caur ",main="Monthly Unemployment Rate in California from 1976 to 2022")
```

```{r}
# Assign more meaningful variable names
colnames(data)<-c("Period","unemployment_rate")
# Convert data into time series dataset
attach(data)
datats<-ts(unemployment_rate  ,c(1976,1),c(2022,9),12)
str(datats)
```



```{r}
datats
```
# Check the frequency of the time series data
```{r}
frequency(datats)
```

```{r}
# Review the summary statistics
summary(datats)
```

```{r}
plot(datats,xlab="Date", ylab = " caur",main="Monthly Unemployment Rate in California from 1976 to 2022")
```
#Boxplot function to see any seasonal effects
```{r}
boxplot(datats~cycle(datats),xlab="Date", ylab = "Caur" ,main ="Monthly Unemployment Rate in California from 1976 to 2022")

```
### From these exploratory plots, we can make some initial inferences:









```{r}
decompose_datats <- decompose(datats,"multiplicative")
autoplot(decompose_datats)
```
### In above decomposed plots we can see the trend and seasonality again but we can also observe the estimation of the random component depicted under the remainder



### Autocorrelation function
```{r}
autoplot(acf(datats,plot=FALSE))+ labs(title="Correlogram of Monthly Unemployment Rate in California from 1976 to 2022")
```
## If the autocorrelation crosses the dashed blue line, it means that specific lag is significantly correlated with current series.



# Fit model for time series


```{r}

```







