---
title: "Appendix B: R Markdown Code"
subtitle: "ADS-506 Final Project"
author: "Nicholas Lee"
date: "12/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## PACKAGES ##
library(tidyverse)
library(fpp2)
library(readr)
library(zoo)
library(tseries) # For adf test
library(rugarch) # For autoarfima()
```

# Importing the Data Set
```{r Import Data}
# Import Dataset #
caur_df <- read_csv(
  "CAUR.csv", 
  col_types = cols(DATE = col_date(format = "%Y-%m-%d")))
```

# Exploratory Data Analysis (EDA)
## Series Summary Statistics
```{r Summary Statistics}
summary(caur_df)
```

## Distribution of the Target Variable
```{r Histogram of CAUR}
hist(caur_df$CAUR,
     main = "Histogram of California Unemployment Rate",
     xlab = "Unemployment Rate %",
     col = "blue")
```

_Note_: The above plot indicates that the distribution of the target variable is skewed right.

## Log + 1 Transformation of the Dependent Variable
```{r}
hist(log(caur_df$CAUR + 1),
     main = "Histogram of Log+1 Transformed California Unemployment Rate",
     xlab = "Unemployment Rate %",
     col = "blue")
```

## Log+1 Transformed Measures of Centrality
```{r}
mean(log(caur_df$CAUR + 1))
median(log(caur_df$CAUR + 1))
```

## Yearly Average Unemployment Rate (%)
```{r CAUR-Yearly Statistics}
# Create a New Yearly Dataframe #
caur_yearly <- caur_df

# Extract Years from DATE Column #
caur_yearly$Year <- format(caur_yearly$DATE, "%Y")

# Drop Original Date Column #
caur_yearly$DATE <- NULL

# Calculate Average Unemployment Rate per Year #
caur_yearly_stats <- caur_yearly %>%
  group_by(Year) %>%
  summarise(Avg_UR = mean(CAUR))

# Plot Yearly Avg. Unemployment Rate #
ggplot(caur_yearly_stats, aes(Year, Avg_UR, group = 1)) +
  geom_line(size = 1, color = "blue") +
  geom_point(color = "blue") + 
  ggtitle("Yearly Average Unemployment Rates (Jan. 1976 - Oct. 2022)") +
  ylab("Unemployment Rate (%)") +
  scale_x_discrete(breaks = round(
    seq(
      min(caur_yearly_stats$Year), 
      max(caur_yearly_stats$Year), by = 5), 1))
```

## Time Plot of California Unemployment Rate
```{r Raw Data - Time Plot}
# Convert Raw Data to Time Series #
caur_ts <- ts(
  caur_df$CAUR, start = c(1976, 1), frequency = 12)

# Time Plot #
autoplot(caur_ts) +
  labs(title = "Monthly Unemployment Rate (Jan. 1976 - Oct. 2022)",
       x = "Time",
       y = "Unemployment Rate (%)") +
  geom_hline(yintercept = mean(caur_df$CAUR), color="blue") +
  theme_minimal()
```

```{r Raw Data - Time Plot Zoomed In}
# Time Plot #
autoplot(caur_ts) +
  labs(title = "Monthly Unemployment Rate (Jan. 1980 - Jan. 1985)",
       x = "Time",
       y = "Unemployment Rate (%)") +
  coord_cartesian(xlim = c(1980, 1985)) +
  theme_minimal()
```

## Distribution of Unemployment Rate by Month
```{r CAUR-Monthly Box Plot}
# Create Column for Month #
caur_monthly <- caur_df 
caur_monthly$Month <- months(caur_df$DATE)

# Drop DATE Column #
caur_monthly$DATE <- NULL

# Reorder Month Names #
caur_monthly$Month <- factor(
  caur_monthly$Month,
  levels = c("January", "February", "March", "April",
             "May", "June", "July", "August",
             "September", "October","November", "December"))

# Boxplot by Month #
ggplot(caur_monthly, aes(reorder(Month, desc(Month)), CAUR)) +
  ggtitle("Distribution of Unemployment Rate by Month") +
  ylab("Unemployment Rate (%)") +
  geom_boxplot() +
  coord_flip()
```

```{r Monthly Outliers}
# Identification of Outliers by Date
caur_df[caur_df$CAUR > 13, ]
```

## Decomposition Plots
```{r Raw Data - Additive Decomposition Plot}
# Decompose Time Series into Components #
caur_decomp <- decompose(caur_ts)

# Decomposition Plot #
plot(caur_decomp)
```

```{r Multiplicative Decomposition}
# Decompose Time Series into Components #
caur_multi_decomp <- decompose(caur_ts, type = "multiplicative")

# Decomposition Plot #
plot(caur_multi_decomp)
```

# Pre-Processing
## Outlier Handling with tsclean()
```{r Outlier Resolution and Time Plot}
# Resolve outliers via tsclean()
caur_ts_nooutlier <- tsclean(caur_ts)

# Time Plot without Outliers
autoplot(caur_ts_nooutlier) +
  labs(title = "Monthly Unemployment Rate - Outliers Removed and Forecasted",
       x = "Time",
       y = "Unemployment Rate (%)")
```

## Decomposition Plot for Series where Outliers are Resolved
```{r Decomposition Plot of Outlier-Resolved Series}
# Decompose Time Series into Components #
caur_nooutlier_decomp <- decompose(
  caur_ts_nooutlier)

# Decomposition Plot #
plot(caur_nooutlier_decomp)
```

# Series Characterization
## Test for Stationarity - Augmented Dickey-Fuller (ADF) Test
```{r ADF Test of Raw Series}
adf.test(caur_ts)
```

|   The null hypothesis of the Augmented Dickey-Fuller (ADF) test is that the series is non-stationary, where the alternative hypothesis is that the series is stationary. The null hypothesis can be rejected if the p-value is less than a significant value, typically 0.05. Here, the p-value is 0.1525, much greater than 0.05. Therefore, the null hypothesis cannot be rejected, indicating the series is non-stationary.


## Assessing if the Series is a Random Walk - Autocorrelation (ACF) Plot
```{r Series ACF Plot}
Acf(caur_ts,
    main = "Autocorrelation Plot")
```

_Note_: According to Shmueli and Lichtendahl (2018, p. 145), a strong positive lag-1 autocorrelation value is indicative of a strong linear trend. Here, the lag-1 autocorrelation is very close to one, indicating the presence of a linear trend in the series. 

|   The slow decrease in autocorrelation values as lag values increase also support the hypothesis that there is a trend in the data. More importantly, the slow decay to zero of the autocorrelation values suggest that the series may be a random walk. If this is the case, forecasting future values will be more challenging than expected.
|   In addition, the lack of any patterns in the ACF plot suggests that there is no seasonality in the series.

## Partial Autocorrelation Plot of the Raw Series
```{r pACF Plot of Raw Series}
Pacf(caur_ts,
     main = "Partial Autocorrelation Plot")
```

|   The partial autocorrelation (pacf) plot suggests that a first or second-order autoregressive model is most suitable for forecasting the series, based on the lags that have partial autocorrelation values over the thresholds. 

## Random Walk Test via AR(1) Slope Coefficient Hypothesis
```{r RandomWalk AR(1) Coef. Test}
# AR(1) Model
ar1_model <- arima(caur_ts, order = c(1,0,0))

# AR(1) Model Summary
summary(ar1_model)
```

|   <mark>**_Warning:_**</mark> The slope coefficient of the AR(1) model is 0.9765. This value is fairly close to one. A random walk results in an AR(1) model with a slope coefficient equal to one. Therefore, this assessment supports the previous finding that the initial series may be a random walk.

## Random Walk Assessment Via an ACF Plot of the Lag-1 Differenced Series
```{r RandomWalk ACF Plot of Differenced Series Test}
# Order-1 Differencing of the Series
caur_ts_diff1 <- diff(caur_ts)

# ACF Plot of Order-1 Differenced Series
Acf(caur_ts_diff1,
    main = "Autocorrelation of Order-1 Differenced Series")
```

_Note_: If the ACF plot for all lags are near zero and within the significant value thresholds (shown above by dashed, horizontal blue lines), then the original series is most likely a random walk (Shmueli & Lichtendahl, 2018). In the above plot, the autocorrelation of the order-1 differenced series at lag-1 is above the threshold values; however, the value for the lag-1 autocorrelation of the differenced series is slightly above 0.2. The above plot does not conclusively indicate that the series is not a random walk.

## Random Walk Assessment via ADF Test of Lag-1 Differenced Series
```{r lag-1 differenced ADF Test}
# ADF Test of Lag-1 Differenced Series
adf.test(caur_ts_diff1)
```

|   The p-value of the ADF test on the lag-1 differenced series is 0.01, lower than the significance threshold of 0.05. This result indicates that the null hypothesis can be rejected in favor of the alternative hypothesis; thus, the differnced series is stationary. If the lag-1 differenced series results in stationarity, the undifferenced series is most likely a random walk (Brownlee, 2017).


## Remove of Trend via lag-1 Differencing
```{r Time plot with lag-1 differencing}
# Time Plot of the Lag-1 Differenced Series
autoplot(caur_ts_diff1) +
  labs(
    y = "Difference",
    main = "Time Plot of Lag-1 Differenced Series"
  )
```

# Data Partitioning
```{r Data Partitioning}
# Use the series where outliers have been resolved
# Training Set: Jan. 1976 to Dec. 2018
caur_train <- window(
  caur_ts_nooutlier,
  start = c(1976, 1),
  end = c(2018, 12)
)

# Test Set: Jan. 2019 - Oct. 2022
# 46 months
caur_test <- window(
  caur_ts_nooutlier,
  start = c(2019, 1)
)
```

# Forecasting Methods
## Naive Forecast Modeling
```{r Naive Forecast}
# rwf(): Naive and Random Walk Forecasting
caur_rwf <- rwf(
  caur_train,
  h = 46
)

autoplot(caur_rwf,
         alpha = 0.3,
         series = "Naive Forecast") +
  autolayer(caur_test,
            series = "Test Partition") +
  labs(
    y = "Unemployment Rate (%)",
    title = "Naive Forecast of Test Partition"
  ) 
```

```{r Naive Forecast Zoomed In}
autoplot(caur_rwf) +
  labs(
    y = "Unemployment Rate (%)",
    title = "Naive Forecast for Nov.'22 Until Dec.'23"
  ) +
  coord_cartesian(xlim = c(2021, 2024))
```

## Double-Exponential Smoothing
```{r Double-Exponential Model and Plot}
# For series with trend but no seasonality
# ACF plots revealed the series has no seasonal components

# Double-Exponential (Holt's) Model
holt_model <- holt(
  caur_train,
  h = 46
)

# Plot Series with Forecasts
autoplot(holt_model, 
         series = "Double-Exponential Forecast",
         alpha = 0.5) +
  autolayer(caur_test, series = "Test Set") +
  labs(
    y = "Unemployment Rate (%)",
    title = "Double-Exponential Forecasts"
  ) 

# summary(holt_model)
# Training Set Metrics:
## RMSE: 0.063360
## AIC: 193.5718

```

```{r Double-Exponential Zoomed In}
autoplot(holt_model, 
         series = "Double-Exponential Forecast",
         alpha = 0.3) +
  autolayer(caur_test, series = "Test Set") +
  labs(
    y = "Unemployment Rate (%)",
    title = "Double-Exponential Forecasts"
  ) + 
  coord_cartesian(xlim = c(2018, 2024))
```

## Holt-Winter's Smoothing (Triple-Exponential Smoothing)
```{r ETS Undifferenced Series}
# Holt-Winter's Model vis ets()
caur_ets <- ets(
  caur_train,
  model = "AAN",
  damped = TRUE,
  alpha = 0.75
)

# summary(caur_ets)
# Automated Parameter Detection
## Optimal Model Identified as: "A, Ad, N"
## Optimal Alpha: 0.75
# Model ANN, varying alphas:
## Alpha = 0.10, AIC = 1,914.83, RMSE = 0.27879
## Alpha = 0.25, AIC = 1,104.42, RMSE = 0.12713
## Alpha = 0.50, AIC = 546.56, RMSE = 0.07404
## Alpha = 0.75, AIC = 382.30, RMSE = 0.06315
## Alpha = 0.90, AIC = 403.30, RMSE = 0.06444
```

### Holt-Winter's Forecast
```{r HoltWinter Model Forecasts and Plot}
# 166 Month Forecast - Nov.'22 - Dec.'23
caur_ets_forecast <- forecast(
  caur_ets,
  h = 46
)

# ETS Forecast Plot
autoplot(caur_train,
          series = "Training Partition") +
  autolayer(caur_ets_forecast,
            series = "Holt-Winter's Model Forecasts",
            alpha = 0.4) +
  autolayer(caur_test,
            series = "Test Partition") +
  labs(
    y = "Unemployment Rate (%)",
    title = "Holt-Winter's Model Forecasts"
  )
```

```{r HoltWinters Forecast Plot Zoomed In}
# ETS Forecast Plot
autoplot(caur_train,
          series = "Training Partition") +
  autolayer(caur_ets_forecast,
            series = "Holt-Winter's Model Forecasts",
            alpha = 0.4) +
  autolayer(caur_test,
            series = "Test Partition") +
  labs(
    y = "Unemployment Rate (%)",
    title = "Holt-Winter's Model Forecasts"
  ) +
  coord_cartesian(xlim = c(2018, 2023))
```

## ARIMA Model
```{r ARIMA Model}
# Automated ARIMA Parameters
caur_autoARIMA <- auto.arima(
  caur_train)
# summary(caur_autoARIMA)
## Automatic parameter selection: ARIMA(2,0,2)(1,0,2)[12]
## AIC: -1,396.23, RMSE: 0.06084

# Manual ARIMA Parameter Search
caur_arima <- Arima(
  caur_train,
  order = c(2,1,4)
)

# summary(caur_arima)
# ARIMA (1,0,1)
## AIC: -890.51, RMSE: 0.10067
# ARIMA (1,1,1)
## AIC: -1,331.36, RMSE: 0.06590
# ARIMA (1,1,2)
## AIC: -1,380.53, RMSE: 0.06268
# ARIMA (2,0,2)
## AIC: -1,382.37, RMSE: 0.06212
# ARIMA (2,1,2)
## AIC: -1,380.23, RMSE: 0.06257
# ARIMA (2,1,3)
## AIC: -1,377.41, RMSE: 0.06263
# ARIMA (2,1,4)
## AIC: -1,381.47, RMSE: 0.06225
```

#### Initial Manual Selection of ARIMA Parameters
* p: 1 (PACF plot revealed 1 lag with significant partial autocorrelations)
* d: 1 (1 order of differencing resulted in stationarity)
* q: 2 (ACF plot showed multiple significant ACF values - use automated selection)

### Various ARIMA Models' Forecasts Comparison
```{r ARIMA Forecast Comparison}
caur_autoARIMA_forecast <- forecast(
  caur_autoARIMA,
  h = 46
)

autoplot(fitted(caur_autoARIMA),
         series = "Training Partition") +
  autolayer(caur_autoARIMA$x,
            colour = TRUE,
            series = "Fitted Model") +
  autolayer(caur_test,
            series = "Test Partition") +
  autolayer(caur_autoARIMA_forecast,
            series = "Automated Model Forecasts",
            alpha = 0.4) +
  labs(
    y = "Unemployment Rate (%)", 
    title = ("Automated ARIMA(2,0,2)(1,0,2)[12] Model Test Set Forecasts")
  )
```

```{r Manual ARIMA Model Fitting and Plot}
caur_arima_forecast <- forecast(
  caur_arima,
  h = 46
)

autoplot(fitted(caur_arima),
         series = "Fitted Model") +
  autolayer(caur_arima$x,
            colour = TRUE,
            series = "Training Partition") +
  autolayer(caur_test,
            series = "Test Partition") +
  autolayer(caur_arima_forecast,
            series = "Manual Model Forecasts",
            alpha = 0.4) +
  labs(
    y = "Unemployment Rate (%)", 
    title = ("Manual ARIMA Model ARIMA(2,1,4) Test Set Forecasts")
  )
```

```{r ARIMA Forecasts Comparisons}
autoplot(caur_test,
          series = "Test Partition") +
  autolayer(caur_autoARIMA_forecast,
            series = "Automated ARIMA Model Forecasts",
            alpha = 0.3) +
  autolayer(caur_arima_forecast,
            series = "Manual ARIMA Model Forecasts",
            alpha = 0.3) +
  labs(
    y = "Unemployment Rate (%)",
    title = "Comparing Automated and Manual ARIMA Model Forecasts"
  )
```

__Note__: The test partition is fully captured within the 95% confidence interval of both ARIMA models. However, more of the data is captured in the 80% confidence interval of the automated model than in 80% confidence interval of the manual model. Thus, moving forward, only the automated model will be considered.

## ARFIMA Model
```{r ARFIMA Model}
caur_autoARFIMA <- autoarfima(
  caur_train,
  ar.max = 2,
  ma.max = 3,
  criterion = "AIC",
  method = "full",
  arfima = TRUE
)

# ARFIMA Forecast
caur_autoARFIMA_forecast <- arfimaforecast(
  caur_autoARFIMA$fit,
  n.ahead = 46
)

caur_autoARFIMA_forecast_ts <- ts(
  caur_autoARFIMA_forecast@forecast$seriesFor,
  start = c(2019, 1),
  frequency = 12)

# ARFIMA Forecast Plot
autoplot(caur_test) +
  autolayer(caur_autoARFIMA_forecast_ts,
            series = "Automated ARFIMA Model Forecasts") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARFIMA Model Test Set Forecasts"
  )

# Model Residuals
# caur_autoARFIMA_forecast@model$modeldata$residuals)
```

## All Models Assessed (Test Partition Time Plot)
```{r All Models Test Partition Time Plot}
# Naive Forecast
caur_naive <- rwf(
  caur_train, h = 46)

autoplot(caur_test,
         series = "Test Partition") +
  autolayer(caur_naive$mean,
            series = "Naive Forecast") +
  autolayer(holt_model$mean,
            series = "Double-Exponential") +
  autolayer(caur_ets_forecast$mean,
            series = "Holt-Winters Model") +
  autolayer(caur_autoARIMA_forecast,
            series = "ARIMA(2,0,2)(1,0,2)[12] Model",
            alpha = 0.3) +
  autolayer(caur_autoARFIMA_forecast_ts,
            series = "ARFIMA Model") +
  labs(
    y = "Unemployment Rate (%)",
    title = "All Models Test Partition Forecast"
  )
```

# Evaluation Metrics
```{r RMSE and MAPE Calculation}
# Calculate Residuals #
# Naive Forecast Residuals #
naive_resid <- caur_test - caur_rwf$mean

# Double-Exponential/Holt's Model Residuals #
holt_resid <- caur_test - holt_model$mean

# Holt-Winters Model Residuals #
ets_resid <- caur_test - caur_ets_forecast$mean

# ARIMA(2,0,2)(1,0,2)[12] Residuals #
arima_resid <- caur_test - caur_autoARIMA_forecast$mean

# ARFIMA Residuals #
arfima_resid <- caur_test - caur_autoARFIMA_forecast_ts

# RMSE = (sqrt(mean(errors^2))) #
rmse <- function(resid_series) {
  calculated_rmse <- sqrt(mean(resid_series^2))
  return(calculated_rmse)
}

## Training Set RMSE ##
# naive_train_rmse <- rmse(caur_rwf$residuals)
holtExp_train_rmse <- rmse(holt_model$residuals)
ets_train_rmse <- rmse(caur_ets$residuals)
arima_train_rmse <- rmse(caur_autoARIMA$residuals)
arfima_train_rmse <- rmse(
  caur_autoARFIMA_forecast@model$modeldata$residuals)

## Test Set RMSE ##
naive_rmse <- rmse(naive_resid)
holt_rmse <- rmse(holt_resid)
ets_rmse <- rmse (ets_resid)
arima_rmse <- rmse(arima_resid)
arfima_rmse <- rmse(arfima_resid)

# MAPE = sum(abs(errors/actual)) * 100 * 1/test_size #
mape <- function(resid_series, actual_series) {
  calcualted_mape <- sum(
    abs(resid_series/actual_series)) * 100 * (1/length(actual_series))
  return(calcualted_mape)
}

## Training Set MAPE ##
# naive_train_mape <- mape(caur_rwf$residuals, caur_train)
holtExp_train_mape <- mape(holt_model$residuals, caur_train)
ets_train_mape <- mape(caur_ets$residuals, caur_train)
arima_train_mape <- mape(caur_autoARIMA$residuals, caur_train)
arfima_train_mape <- mape(
  caur_autoARFIMA_forecast@model$modeldata$residuals, caur_train)


## Test Set MAPE ##
naive_mape <- mape(naive_resid, caur_test)
holt_mape <- mape(holt_resid, caur_test)
ets_mape <- mape(ets_resid, caur_test)
arima_mape <- mape(arima_resid, caur_test)
arfima_mape <- mape(arfima_resid, caur_test)


# Evaluation Metrics Table #
model_eval_metrics <- data.frame(
  Models = c("Naive", "Double-Exponential Smoothing", "Holt-Winters",
             "ARIMA(2,0,2)(1,0,2)[12]", "ARFIMA"),
  Train_RMSE = c(0.1286977, holtExp_train_rmse, ets_train_rmse, 
                 arima_train_rmse, arfima_train_rmse),
  Test_RMSE = c(naive_rmse, holt_rmse, ets_rmse, 
                arima_rmse, arfima_rmse),
  Train_MAPE = c(1.222649, holtExp_train_mape, ets_train_mape, 
                 arima_train_mape, arfima_train_mape),
  Test_MAPE = c(naive_mape, holt_mape, ets_mape, 
                arima_mape, arfima_mape)
)

model_eval_metrics
```

# Final ARFIMA Model
```{r Final Model: Fitting and Forecast}
# Final Model Forecast #
caur_fulldata_arfima <- autoarfima(
  caur_ts,
  ar.max = 2,
  ma.max = 3,
  criterion = "AIC",
  method = "full",
  arfima = TRUE
)

# ARFIMA Full Data Forecast
caur_fulldata_ARFIMA_forecast <- arfimaforecast(
  caur_fulldata_arfima$fit,
  n.ahead = 14
)

# Convert forecast to time series
caur_fulldata_ARFIMA_forecast_ts <- ts(
  caur_fulldata_ARFIMA_forecast@forecast$seriesFor,
  start = c(2022, 11),
  frequency = 12)

# ARFIMA Forecast Plot
autoplot(caur_ts) +
  autolayer(caur_fulldata_ARFIMA_forecast_ts,
            series = "ARFIMA Model Forecasts") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARFIMA Model Forecasts for Nov.'22 - Dec.'23"
  )
```

```{r Final Forecast Plot Zoomed In}
# ARFIMA Forecast Plot
autoplot(caur_ts) +
  autolayer(caur_fulldata_ARFIMA_forecast_ts,
            series = "ARFIMA Model Forecasts") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARFIMA Model Forecasts for Nov.'22 - Dec.'23"
  ) +
  coord_cartesian(xlim = c(2022, 2024))
```

```{r ARFIMA Forecats Values}
caur_fulldata_ARFIMA_forecast_ts
```


# Seasonal-ARIMA Future Forecasts
```{r Seasonal ARIMA Future Forecasts}
# Create ARIMA(2,0,2)(1,0,2)[12] Model
caur_seasonal_arima <- Arima(
  caur_ts,
  order = c(2,0,2),
  seasonal = c(1,0,2)
)

# Forecast 14 Months into the future
caur_seasonal_arima_forecast <- forecast(
  caur_seasonal_arima,
  h = 14
)

# Plot Future Forecast
autoplot(caur_ts) +
  autolayer(caur_seasonal_arima_forecast,
            series = "2023 Forecast") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARIMA(2,0,2)(1,0,2) Forecast Nov.'22 - Dec.'23"
  )
```

```{r Seasonal ARIMA Forecast Plot Zoomed In}
# Plot Future Forecast
autoplot(caur_ts) +
  autolayer(caur_seasonal_arima_forecast,
            series = "2023 Forecast") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARIMA(2,0,2)(1,0,2) Forecast Nov.'22 - Dec.'23"
  ) +
  coord_cartesian(xlim = c(2021,2024))
```

# Future Forecasts - AFRIMA and ARIMA Overlay
```{r}
autoplot(caur_ts) +
  autolayer(caur_seasonal_arima_forecast,
            alpha = 0.3,
            series = "2023 ARIMA Forecast") +
  autolayer(caur_fulldata_ARFIMA_forecast_ts,
            series = "ARFIMA Model Forecasts") +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARIMA and ARFIMA Future Forecast (Nov.'22 - Dec,'23) Overlay"
  ) +
  coord_cartesian(xlim = c(2021,2024))
```

## Visualizations for Presentation 
```{r Line Plot: Colored by Recession Periods - Future Forecasts}
autoplot(caur_ts) +
  autolayer(caur_seasonal_arima_forecast,
            alpha = 0.3,
            series = "2023 ARIMA Forecast") +
  autolayer(caur_fulldata_ARFIMA_forecast_ts,
            series = "ARFIMA Model Forecasts") +
  annotate("rect", fill = "grey", alpha = 0.5, 
        xmin = 1980, xmax = 1980 + 7/12,
        ymin = -Inf, ymax = Inf) +
   annotate("rect", fill = "grey", alpha = 0.5, 
        xmin = 1981 + 7/12, xmax = 1982 + 11/12,
        ymin = -Inf, ymax = Inf) +
   annotate("rect", fill = "grey", alpha = 0.5, 
        xmin = 1990 + 7/12, xmax = 1991 + 3/12,
        ymin = -Inf, ymax = Inf) +
   annotate("rect", fill = "grey", alpha = 0.5, 
        xmin = 2001 + 3/12, xmax = 2001 + 11/12,
        ymin = -Inf, ymax = Inf) +
   annotate("rect", fill = "grey", alpha = 0.5, 
        xmin = 2007 + 7/12, xmax = 2009 + 6/12,
        ymin = -Inf, ymax = Inf) +
  labs(
    y = "Unemployment Rate (%)",
    title = "ARIMA and ARFIMA Future Forecast (Nov.'22 - Dec,'23) Overlay"
  ) +
  theme(legend.position = "bottom") +
  theme_bw()
```

_Note_: Recessions occurred:
* Jan. 1980 - July 1980
* July 1981 - Nov. 1982
* July 1990 - Mar. 1991
* Mar. 2001 - Nov. 2001
* Dec. 2007 - June 2009

```{r Line Plot: Raw Series with Recession Periods}
autoplot(caur_ts) +
  annotate("rect", fill = "red", alpha = 0.5, 
        xmin = 1980, xmax = 1980 + 7/12,
        ymin = -Inf, ymax = Inf) +
  annotate("rect", fill = "red", alpha = 0.5, 
        xmin = 1981 + 7/12, xmax = 1982 + 11/12,
        ymin = -Inf, ymax = Inf) +
  annotate("rect", fill = "red", alpha = 0.5, 
        xmin = 1990 + 7/12, xmax = 1991 + 3/12,
        ymin = -Inf, ymax = Inf) +
  annotate("rect", fill = "red", alpha = 0.5, 
        xmin = 2001 + 3/12, xmax = 2001 + 11/12,
        ymin = -Inf, ymax = Inf) +
  annotate("rect", fill = "red", alpha = 0.5, 
        xmin = 2007 + 7/12, xmax = 2009 + 6/12,
        ymin = -Inf, ymax = Inf) +
  labs(
    y = "Unemployment Rate (%)",
    title = "California Unemployment Rates Jan. 1976 - Oct. 2022"
  ) +
  theme(legend.position = "bottom") +
  theme_bw()
```

```{r}
autoplot(caur_ts) +
  autolayer(forecast(rwf(caur_ts, h = 14))$mean,
            series = "naive",
            lty = 2,
            lwd = 1) +
  autolayer(holt(caur_ts, h = 14)$mean,
            series = "Double-Exponential",
            lty = 2,
            lwd = 1) +
  autolayer(forecast(ets(
    caur_ts, model = "AAN", damped = TRUE, alpha = 0.75), h = 14)$mean,
    series = "Holt-Winter's",
    lty = 2,
    lwd = 1) +
  autolayer(caur_seasonal_arima_forecast$mean,
            series = "ARIMA",
            lty = 2,
            lwd = 1) +
  autolayer(caur_fulldata_ARFIMA_forecast_ts,
            series = "ARFIMA",
            lty = 2,
            lwd = 1) +
  labs(
    y = "Unemployment Rate (%)",
    title = "Expected Future Unemployment Rates"
  ) +
  coord_cartesian(xlim = c(2009, 2024)) +
  theme(legend.position = "bottom")
```



**References**

Brownlee, J. (2017). _A gentle introduction to the random walk for times series forecasting with python_. Machine Learning Mastery. Retrieved November 15, 2022, from https://machinelearningmastery.com/gentle-introduction-random-walk-times-series-forecasting-python/ 

Hamilton, J. D. (n.d.). _The Econbrowser Recession Indicator Index_. Econbrowser. Retrieved December 1, 2022, from https://econbrowser.com/recession-index 


Shmueli, G., & Lichtendahl Jr, K. C. (2018). _Practical time series forecasting with r: A hands-on guide_. Axelrod Schnall Publishers.