---
title: "ADS-506 Final Project"
author: "Nicholas Lee"
date: "11/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
## PACKAGES ##
library(tidyverse)
library(fpp2)
library(readr)
library(zoo)
library(tseries) # For adf test
```

# Importing the Data Set
```{r Import Data}
# Import Dataset #
caur_df <- read_csv(
  "CAUR.csv", 
  col_types = cols(DATE = col_date(format = "%Y-%m-%d")))
```

# Series Summary Statistics
```{r Summary Statistics}
summary(caur_df)
```

```{r Histogram of CAUR}
hist(caur_df$CAUR,
     main = "Histogram of California Unemployment Rate",
     xlab = "Unemployment Rate %",
     col = "blue")
```

_Note_: The above plot indicates that the distribution of the target variable is skewed right.

# Yearly Average Unemployment Rate (%)
```{r CAUR-Yearly Statistics}
# Create a New Yearly Dataframe #
caur_yearly <- caur_df

# Extract Years from DATE Column #
caur_yearly$Year <- format(caur_yearly$DATE, "%Y")

# Drop Original Date Column #
caur_yearly$DATE <- NULL

# Calculate Average Unemployment Rate per Year #
caur_yearly_stats <- caur_yearly %>%
  group_by(Year) %>%
  summarise(Avg_UR = mean(CAUR))

# Plot Yearly Avg. Unemployment Rate #
ggplot(caur_yearly_stats, aes(Year, Avg_UR, group = 1)) +
  geom_line(size = 1, color = "blue") +
  geom_point(color = "blue") + 
  ggtitle("Yearly Average Unemployment Rates (Jan. 1976 - Sept. 2022)") +
  ylab("Unemployment Rate (%)") +
  scale_x_discrete(breaks = round(
    seq(
      min(caur_yearly_stats$Year), 
      max(caur_yearly_stats$Year), by = 5), 1))
  theme_minimal()
```

# Time Plot of California Unemployment Rate
```{r Raw Data - Time Plot}
# Convert Raw Data to Time Series #
caur_ts <- ts(
  caur_df$CAUR, start = c(1976, 1), frequency = 12)

# Time Plot #
autoplot(caur_ts) +
  labs(title = "Monthly Unemployment Rate (Jan. 1976 - Sept. 2022)",
       x = "Time",
       y = "Unemployment Rate (%)") +
  geom_hline(yintercept = mean(caur_df$CAUR), color="blue") +
  theme_minimal()
```

```{r Raw Data - Time Plot Zoomed In}
# Time Plot #
autoplot(caur_ts) +
  labs(title = "Monthly Unemployment Rate (Jan. 1980 - Jan. 1985)",
       x = "Time",
       y = "Unemployment Rate (%)") +
  coord_cartesian(xlim = c(1980, 1985)) +
  theme_minimal()
```

# Distribution of Unemployment Rate by Month
```{r CAUR-Monthly Box Plot}
# Create Column for Month #
caur_monthly <- caur_df 
caur_monthly$Month <- months(caur_df$DATE)

# Drop DATE Column #
caur_monthly$DATE <- NULL

# Reorder Month Names #
caur_monthly$Month <- factor(
  caur_monthly$Month,
  levels = c("January", "February", "March", "April",
             "May", "June", "July", "August",
             "September", "October","November", "December"))

# Boxplot by Month #
ggplot(caur_monthly, aes(Month, CAUR)) +
  ggtitle("Distribution of Unemployment Rate by Month") +
  ylab("Unemployment Rate (%)") +
  geom_boxplot()
```

```{r Monthly Outliers}
caur_df[caur_df$CAUR > 13, ]
```

# Decomposition Plots
```{r Raw Data - Additive Decomposition Plot}
# Decompose Time Series into Components #
caur_decomp <- decompose(caur_ts)

# Decomposition Plot #
plot(caur_decomp)
```

```{r Multiplicative Decomposition}
# Decompose Time Series into Components #
caur_multi_decomp <- decompose(caur_ts, type = "multiplicative")

# Decomposition Plot #
plot(caur_multi_decomp)
```

# Test for Stationarity - Augmented Dickey-Fuller (ADF) Test
```{r}
adf.test(caur_ts)
```

|   The null hypothesis of the Augmented Dickey-Fuller (ADF) test is that the series is non-stationary, where the alternative hypothesis is that the series is stationary. The null hypothesis can be rejected if the p-value is less than a significant value, typically 0.05. Here, the p-value is 0.1525, much greater than 0.05. Therefore, the null hypothesis cannot be rejected, indicating the series is non-stationary.


# Assessing if the Series is a Random Walk - Autocorrelation (ACF) Plot
```{r Series ACF Plot}
Acf(caur_ts,
    main = "Autocorrelation Plot")
```

_Note_: According to Shmueli and Lichtendahl (2018, p. 145), a strong positive lag-1 autocorrelation value is indicative of a strong linear trend. Here, the lag-1 autocorrelation is very close to one, indicating the presence of a linear trend in the series. 

|   The slow decrease in autocorrelation values as lag values increase also support the hypothesis that there is a trend in the data. More importantly, the slow decay to zero of the autocorrelation values suggest that the series may be a random walk. If this is the case, forecasting future values will be more challenging than expected.
|   In addition, the lack of any patterns in the ACF plot suggests that there is no seasonality in the series.

# Random Walk Assessment Via an ACF Plot
```{r RandomWalk ACF Plot Test}
# Order-1 Differencing of the Series
caur_ts_diff1 <- diff(caur_ts)

# ACF Plot of Order-1 Differenced Series
Acf(caur_ts_diff1,
    main = "Autocorrelation of Order-1 Differenced Series")
```

_Note_: If the ACF plot for all lags are near zero and within the significant value thresholds (shown above by dashed, horizontal blue lines), then the original series is most likely a random walk (Shmueli & Lichtendahl, 2018). In the above plot, the autocorrelation of the order-1 differenced series at lag-1 is above the threshold values; however, the value for the lag-1 autocorrelation of the differenced series is slightly above 0.2. The above plot does not conclusively indicate that the series is not a random walk.

# Random Walk Test via AR(1) Slope Coefficient Hypothesis
```{r RandomWalk AR(1) Coef. Test}
# AR(1) Model
ar1_model <- arima(caur_ts, order = c(1,0,0))

# AR(1) Model Summary
summary(ar1_model)
```

|   <mark>**_Warning:_**</mark> The slope coefficient of the AR(1) model is 0.9765. This value is fairly close to one. A random walk results in an AR(1) model with a slope coefficient equal to one. Therefore, this assessment supports the previous finding that the initial series may be a random walk.

# Remove of Trend via lag-1 Differencing
```{r}
# Time Plot of the Lag-1 Differenced Series
autoplot(caur_ts_diff1) +
  labs(
    y = "Difference",
    main = "Time Plot of Lag-1 Differenced Series"
  )
```



**References**

Shmueli, G., & Lichtendahl Jr, K. C. (2018). _Practical time series forecasting with r: A hands-on guide_. Axelrod Schnall Publishers.